The Power of Dependent Reflection

### 2021-07-01 paper.tex
Drafting new abstract.

%% Title information
\title[]{Extraction once again}

@@

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  Dependently-typed host languages empower users to verify a wide
  range of properties of embedded languages and
  programs written in them.  Designers of such embedded languages
  are faced with a difficult choice between using a shallow or a deep
  embedding.  The former is easier to use because the entire
  infrastructure of the host langauge is immediately
  available. Meanwhile, the latter gives full access to the
  structure of embedded programs, but is difficult to use in practice,
  especially when the embedded language is itself dependently typed.

  The main insight presented in this paper is that the choice between
  shallow and deep embedding can be eliminated by working in a host
  language with reflection capabilities: we start from a shallow
  embedding that can use all libraries and tools of the host language,
  and later use reflection to expose the deep structure of the
  embedded programs.
  %
  Concretely, we apply this technique to embed three
  programming languages --- Kaleidoscope, SaC, and (a subset of) APL ---
  into the dependently typed theorem prover Agda, using dependent
  types to statically enforce several properties of interest.  We then
  use Agda's reflection capabilities to extract the embedded programs
  back into the original language, so that the existing toolchain can
  be leveraged.
  %
  In this process, statically verified properties of the host language
  are mapped onto runtime checks in the target language, allowing
  extracted programs to interact safely with existing code.
  %
  Finally, we demonstrate the feasibility of our approach with the
  implementation and extraction of a convolutional neural network in
  our embedding of APL.\@

>>>>

%% Title information
\title[]{The Power of Dependent Reflection}

@@

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}

  In this paper we explore a way to write correct by construction
  programs in languages that do not have built-in verification
  capabilities.  This is achieved in three steps.  Firstly, we
  shallowly embed the language (or its subset) inside of a
  dependently-typed system.  Secondly, we write a program
  within that embedding using dependent types to guarantee
  any property of interest.  Thirdly, we provide a way to
  translate the embedded program back into the original language,
  \eg{} so that it can be used by the existing compilers.

  The main focus of the paper is the third step, which we
  call \emph{extraction}.  The main insight of the paper is
  the use of reflection mechanism of a dependently-typed language
  to achieve extraction.  As a result, the embedding, the
  program and the extractor are written in a single language.

  To demonstrate these ideas precisely we embed a subset of
  the PostScript language in Agda and explain all the three
  steps.

%  Dependently-typed host languages empower users to verify a wide
%  range of properties of embedded languages and
%  programs written in them.  Designers of such embedded languages
%  are faced with a difficult choice between using a shallow or a deep
%  embedding.  The former is easier to use because the entire
%  infrastructure of the host langauge is immediately
%  available. Meanwhile, the latter gives full access to the
%  structure of embedded programs, but is difficult to use in practice,
%  especially when the embedded language is itself dependently typed.
%
%  The main insight presented in this paper is that the choice between
%  shallow and deep embedding can be eliminated by working in a host
%  language with reflection capabilities: we start from a shallow
%  embedding that can use all libraries and tools of the host language,
%  and later use reflection to expose the deep structure of the
%  embedded programs.
%  %
%  Concretely, we apply this technique to embed three
%  programming languages --- Kaleidoscope, SaC, and (a subset of) APL ---
%  into the dependently typed theorem prover Agda, using dependent
%  types to statically enforce several properties of interest.  We then
%  use Agda's reflection capabilities to extract the embedded programs
%  back into the original language, so that the existing toolchain can
%  be leveraged.
%  %
%  In this process, statically verified properties of the host language
%  are mapped onto runtime checks in the target language, allowing
%  extracted programs to interact safely with existing code.
%  %
%  Finally, we demonstrate the feasibility of our approach with the
%  implementation and extraction of a convolutional neural network in
%  our embedding of APL.\@
